Gut microbes and their genes or something else

### 2023-02-17 manuscript/tex/sections/intro.tex
start working on refs

The gut and the brain are intimately linked. Signals from the brain
reach the gut through the autonomic nervous system and the endocrine
system, and the gut can communicate with the brain through the vagus
nerve and through endocrine and immune (cytokine) signaling molecules
\href{https://www.zotero.org/google-docs/?i3SRPn}{(Cerdó et al., 2019;
Pronovost and Hsiao, 2019; Sharon et al., 2016; Tognini, 2017)}. In
addition, the products of microbial metabolism generated in the gut can
influence the brain, both indirectly by stimulating the enteric nervous
and immune systems, as well as directly through molecules that enter
circulation and cross the blood-brain barrier. Causal links between the
gut microbiome and neural development, particularly atypical
development, are increasingly being identified
\href{https://www.zotero.org/google-docs/?UsrFKB}{(Spichak et al.,
2021)}. Both human epidemiology and animal models point to the effects
of gut microbes on the development of autism spectrum disorder
\href{https://www.zotero.org/google-docs/?Y8l5A0}{(Laue et al., 2020;
Wan et al., 2021)} and particular microbial taxa have been associated
with depression
\href{https://www.zotero.org/google-docs/?mbaFE3}{(Mayneris-Perxachs et
al., 2022; Valles-Colomer et al., 2019)} and Alzheimer's disease
\href{https://www.zotero.org/google-docs/?0BEGUn}{(Fung et al., 2017;
Kim et al., 2021)}. But information about this ``microbiome-gut-brain
axis'' in normal neurocognitive development remains lacking,
particularly early in life.

>>>>

The gut and the brain are intimately linked. Signals from the brain
reach the gut through the autonomic nervous system and the endocrine
system, and the gut can communicate with the brain through the vagus
nerve and through endocrine and immune (cytokine) signaling molecules
\cite{cerdoEarlyNutritionGut2019,pronovostPerinatalInteractionsMicrobiome2019,sharonCentralNervousSystem2016,togniniGutMicrobiotaPotential2017}
In addition, the products of microbial metabolism generated in the gut can
influence the brain, both indirectly by stimulating the enteric nervous
and immune systems, as well as directly through molecules that enter
circulation and cross the blood-brain barrier. Causal links between the
gut microbiome and neural development, particularly atypical
development, are increasingly being identified \cite{spichakMiningMicrobesMental2021}.
Both human epidemiology and animal models point to the effects
of gut microbes on the development of autism spectrum disorder
\cite{laueProspectiveAssociationsInfant2020,wanUnderdevelopmentGutMicrobiota2021}
and particular microbial taxa have been associated with depression
\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,valles-colomerNeuroactivePotentialHuman2019}
and Alzheimer's disease \cite{fungInteractionsMicrobiotaImmune2017,kimProbioticSupplementationImproves2021}.
But information about this "microbiome-gut-brain
axis" in normal neurocognitive development remains lacking,
particularly early in life.
